Efficiently stable presentations from error-correcting codes

### 2025-10-16 efficient_stability.tex
Updates from Overleaf

\begin{remark}\label{rk-mus}
Fixing a distribution $\mu_R$ on $R$ only, there is a natural induced distribution $\mu_S$ on $S$ which is obtained by sampling $r\in R$ from $\mu_R$, and then sampling an $s\in S$ according to the atomic measure induced by $R$ (seen as a multiset). For example, if $r=aba$ then conditioned on having chosen $r$, $a$ is given probability $2/3$ and $b$ is given probability $1/3$. Throughout this text, $\mu_S$ is always chosen as this induced distribution and so we often omit it from the notation.
\end{remark}

Definition~\ref{def:eff-stab} specifies what it means for a presentation to be stable, but not when the presentation is \emph{efficiently} stable. 
As mentioned in the introduction, the goal is to optimise the tradeoff between the encoding length of the presentation and the resulting modulus of stability. We now choose a complexity measurement for presentations:
\begin{definition}[Length of a presentation]\label{def:length_of_pres}
    Let $G=\langle S\colon R\rangle$ be a finite presentation. The \emph{length} of the presentation  will be  $$\ell(G)=|S|+ \sum_{r\in R}|r|,$$ where $|r|$ is the length of the word $r\in \mF(S)$ written in the basis $S$.\footnote{Another popular notion of length is the bit length of the encoding of $\langle S\colon R\rangle$, where exponents in $r\in R$ can be written in binary. Up to a factor of $\log |S|$, our notion is stricter than that. See \cite{babai1997short}.}
\end{definition}

@@

As was shortly reviewed in the introduction, and specifically in the related works subsection, many results about various notions of stability are known.  Here, we give two results that are most relevant to our work. First of all, 
for a finite group $G$ we can always write $G=\langle S:R\rangle$ where $S = G$ and $R=\{ g\cdot h \cdot (gh)^{-1} =e \}$. We refer to this presentation as the \emph{multiplication table presentation}\tnote{was ``exhaustive''; edited terminology}. If we let $\mu_S$ and $\mu_R$ be the uniform distribution on $S$ and on $R$ respectively then Definition~\ref{def:eff-stab} reduces to a widely used notion of \emph{flexible (Hilbert-Schmidt) stability}. In particular, for finite groups the following result is known~\cite{gowers2017inverse,de2019operator}. We adopt the formulation from~\cite[Theorem 1.4]{de2022spectral}.

@@

@@

\begin{equation}\label{eq:z2-efficient}
 \Z_2^k = \langle x_1,\ldots,x_k : [x_i,x_j]=e, x_i^2=e \; \forall i\neq j \rangle\;,
\end{equation}
where $[x_i,x_j]=x_ix_jx_i^{-1}x_j^{-1}$ is the group commutator. 
Its length is $O(k^2)$, which is polylogarithmic in the group size instead compared to the polynomial length of the multiplication table presentation. By \cite{hadwin2018stability}, every presentation of a finite group has \textbf{some} modulus of stability.
The following two lemmas show that, though the presentation \eqref{eq:z2-efficient} has a linear modulus of stability, it deteriorates by a constant factor $k$. In some sense, these two examples are on opposite sides of the efficient stability tradeoof: \eqref{eq:z2-efficient} is short but has a bad modulus of stability, while the multiplication table presentation is very long, while having an essentially optimal modulus of stability. As we keep recalling, the goal of this paper is to provide an example of a somewhat short presentation of $\Z_2^k$, which has a good enough modulus of stability to deduce the main technical results needed for MIP*=RE \cite{ji2020mip}.

@@

As one would expect, the dependence of $\delta_k$ on $k$ depends on both the diameter of $\Z_2^k$ for the presentation~\eqref{eq:z2-efficient}, and certain values of its \emph{Dehn function} --- namely, the minimal volume of a Van Kampen diagram with perimeters of length $3$. By expressing each element of $\Z_2^k$ as a product of generators in the natural way, and by applying Corollary~\ref{cor:lin-test}, it is possible to show that $\delta_k=O(k^2\eps)$ in Lemma~\ref{lem:eff-z2}. With more work one can get $\delta_k=O(k\eps)$, see~\cite[Theorem 3.2]{chao2017overlapping}. This turns up to be  tight, as the next lemma shows.

@@

Lemma~\ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k} implies that the modulus of stability of \eqref{eq:z2-efficient} is $\Omega(k\eps)$ whenever $\eps\leq\frac{1}{k-1}$. This is because, for any $0<\eps<\nicefrac{1}{k-1}$, one can add trivial representations of $\Z_2^k$ to copies of the approximate representation described in Lemma~\ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k} (choosing $c=\nicefrac{k}{2}$), such that the resulting map is an $(\eps,\mu_R)$-approximate representation which is at least $(\nicefrac{(k-1)\eps}{32},\mu_S)$-away from any actual representation of $\Z_2^k$. We prove this Lemma in Appendix \ref{appendix:lower_bounds}, which also includes some discussion on the $L^\infty$ variant of it.

@@

\begin{proof}
This follows from Lemma~\ref{lem:eff-z2} and Theorem~\ref{thm:dls-gap}, in a similar manner to our proof sketch of Corollary \ref{cor:Pauli-brading_is_stable}.
\end{proof}

@@

\begin{proof}
The group defined by the right-hand side is obviously abelian and a $2$-group, so it is of the form $\Z_2^{k'}$ for some $k'$. In fact, it is equal to the quotient of $\Z_2^n$ by the subgroup generated by the $\prod_{1\leq j \leq n} x_j^{h_{ij}}$. So it is $\Z_2^k$ where $k=n-\dim\textrm{im}\ h = \dim\ker h=\dim \code$.  
\end{proof}

@@

\begin{definition}[$r$-local linear tester]\label{def:code-test}
Let $\code$ be an $[n,k,d]_q$ linear code and $r\in \N$.
An \emph{$r$-local linear tester for $\code$} is a pair $M = (h,\nu)$ where $h \in \F^{m \times n}$ is a parity check matrix for $\code$, whose every row has Hamming weight at most $r$, and $\nu$ is a distribution over $\{1,\ldots,m\}$. 
\end{definition}

@@

A parity check matrix $h_{\RM} \in \F^{S \times q^m}$ for $\code_{\RM}$, where $S = q^m \times m \times q$, is as follows. Identify the rows with triples $(u,j,t) \in \F_q^m \times \{1,\ldots,m\} \times \F_q$. The $(u,j,t)$-th row of $h_{\RM}$ is the vector in $\F_q^m$ that for $i \in \{0,\ldots,d\}$ has the value $\alpha_{u,v,i}$ for $v = u + t e_j$ in the coordinate indexed by $u + t_i e_j$, the value $-1$ in the coordinate indexed by $u + te_j$, and the value $0$ everywhere else.

@@

 The Reed-Muller code from the previous section is defined over $\F_q$, for $q$ a prime power. We can transform any $q$-ary code, for $q=2^t$, into a binary code using the idea of \emph{code composition} which we now describe.

@@

Let $\kappa: \F_q \to \F_2^t$ denote an invertible linear map such that $\kappa(a)$ is the $\F_2$-representation of $a \in \F_q$ over some (implicitly specified) self-dual basis of $\F_q$ over $\F_2$. We extend $\kappa$ and its inverse $\kappa^{-1}$ to vectors over $\F_q$ coordinate-wise. We let $\tr(\cdot):\F_q\to\F_2$ denote the trace over $\F_2$. Because we chose a self-dual basis for the binary representation, the trace satisfies $\tr(ab)=\kappa(a)\cdot\kappa(b)$  where the right-hand side means the $\F_2$-inner product of the vectors $\kappa(a), \kappa(b)$. For more details on the map $\kappa$ and its properties, see~\cite[Section 3.3]{ji2020mip}.

Let $q=2^t$ and $\code$ be an $[n,k,d]_q$ linear code. Let $\code_{\Had}$ be the Hadamard code over $\F_2^t$ (introduced at the end of Section~\ref{sec:pres-code}). 
Let $\code'$ be the $[qn,tk,d']$ linear code over $\F_2$ defined as follows. Given $a\in (\F_2^t)^{k}$, first map $a\mapsto a'=\kappa^{-1}(a) \in \F_q^{k}$. Then encode $a'$ to $b'=\code(a')\in \F_q^n$. Finally, return $b=\code_\Had(\kappa(b'))\in(\F_2^q)^n$, where $\code_\Had$ is applied component-wise. Using that $\code_\Had$ has distance $q/2$, it is easy to verify that this code has distance $d'\geq dq/2$.

Given an $r$-local tester $M=(h,\nu)$ for $\code$, there is a natural $\max(r,3)$-local tester $M'$ for $\code'$ which can be described as follows. Index coordinates of $\code'$ by pairs $(i,\alpha)\in [n]\times\F_2^t$, fixing a bijection between $[qn]$ and $[n]\times \F_2^t$.  We describe an $\max(r,3)$-local tester $M' = (h',\nu')$ for $\code'$. Informally, $h'$ contains two type of checks. First, the $A$-checks consist of the repetition of $n$ copies of the checks for the Hadamard code, one for each Hadamard-code encoding of an $\F_q$-symbol from $\code$. Second, the $B$-checks implement the checks of $\code$ specified by $M$, directly on the Hadamard encoding. More precisely, define $h'$ to be the block matrix $h'=\begin{pmatrix} A \\ B \end{pmatrix}$ where 
\begin{itemize}
	\item $A \in \F^{nq^2 \times nq}$ is itself a block-diagonal matrix where the diagonal blocks are the $q^2 \times q$ parity check matrix for the Hadamard code. 
	In other words, $A$ can be viewed as $I_{n\times n} \otimes h_\had$ where $h_\had \in \F_2^{q^2 \times q}$ is the parity check matrix for the Hadamard code. 
	\item $B \in \F^{\ell q \times nq}$ is viewed as having rows indexed by pairs $(p,\gamma) \in \{1,\ldots,\ell\} \times \F_q$ and columns indexed by pairs $(i,x) \in \{1,\ldots,n\} \times \F_2^t$. The entry in row $(p,\gamma)$ and column $(i,x)$ is $1$ if and only if $h_{pi} \neq 0$ and $x = \kappa(\gamma h_{pi})$.
\end{itemize}
Define the distribution $\nu'$ as the uniform mixture of the uniform distribution on the rows of the $A$ block matrix and the uniform distribution on the rows of the $B$ block matrix.

>>>>

While the quantitative aspects of this definition of stability depend on the choice of presentation, qualitatively a finitely presented group $G$ is stable with respect to some presentation if and only if it is stable with respect to any presentation. This is shown in Lemma~\ref{lem:stab-group} below, which includes some rough quantitative estimates. We start with a simple but useful calculation.
\tnote{Added the lemma. The proof is pretty annoying. Michael, did you say there are references for it?}

For a set of generators $S$ and an element $r\in\mF(S)$, we let $|r|$ denote the length of the word $r\in \mF(S)$ written in the basis $S\cup S^{-1}$.\footnote{Another popular notion of length is the bit length of the encoding of $\langle S\colon R\rangle$, where exponents in $r\in R$ can be written in binary. Up to a factor of $\log |S|$, our notion is stricter than that. See \cite{babai1997short}.}\tnote{previously, we had just $S$. But we also need to allow inverses, right?}

\begin{lemma}\label{lem:g-phi}
Let $G=\langle S:R\rangle$ and $\mu_S$ a distribution on $S$.  Let $\phi:\mathcal{F}(S)\to \mathcal{U}(\cM)$ and $g:G\to \mathcal{U}(\mN)$ a representation such that $g$ is $(\delta,\mu_S)$ close to $\phi$. Then for any $r\in\mathcal{F}(S)$, it holds that 
\[ \|\phi(r)-w^*g(r)w\|_\tau \,\leq\, \sum_{i=1}^{|r|}  \| w^* g(r_i)w -\phi(r_i)\|_\tau + 2\delta\;,\]
where $r_i$ denotes the $i$-the element of $r$ (which is an element of $S\cup S^{-1}$).
\end{lemma}

\begin{proof}
  Let $r$ be a word in the elements of $S$ and their inverses. For $0\leq i \leq |r|$ we write $r_{\leq i}$ for the length-$i$ prefix of $r$ (which is $1_G$ if $i=0$), and $r_{> i}$ for its length-$(|r|-i)$ suffix (which is $1_G$ if $i=|r|$). 
  Using the triangle inequality, 
  \begin{align*}  
    \|\phi(r)-w^*g(r)w\|_\tau &\leq \sum_{i=1}^{|r|} \| w^*{g}(r_{\leq i}) w{\phi}(r_{>i})-w^* {g}(r_{<i})w\phi(r_{\geq i}))\|_\tau  + \|(w^*w-I_\cM)\phi(r)\|_\tau \\
    &\leq \sum_{i=1}^{|r|} \| w^*g(r_{<i})w(w^* g(r_i)w -\phi(r_i))\phi(r_{>i})\|_\tau \\
    &\qquad + \|w^* g(r_{<i}) (ww^*-P) g(r_i)w \phi(r_{>i})\|_{\tau^\mN} + \|w^*w-I_\cM\|_\tau \\ 
    &\leq \sum_{i=1}^{|r|}\| w^* g(r_i)w -\phi(r_i)\|_\tau + \|ww^*-P\|_{\tau^{\mN}} + \|w^*w-I_\mM\|_\tau\;.
    \end{align*}
    Here, for the last inequality we used that $g(r_i)$, $\phi(r_i)$ are unitaries, and $w$ an isometry, so either of them contracts the norm.  
\end{proof}

\begin{lemma}\label{lem:stab-group}
Suppose that $G=\langle S:R\rangle$ is $(\delta,\mu_S,\mu_R,\mC)$-stable as in Definition~\ref{def:eff-stab}. Let $G = \langle S':R'\rangle$ be another presentation. Then for any presentation $G=\langle S':R'\rangle$, and any distribution $\mu_{R'}$ on $R'$ that has full support, $G=\langle S':R'\rangle$ is $(\delta',\mu_{S'},\mu_{R'},\mC)$-stable for any distribution $\mu_{S'}$ and a modulus $\delta'$ that satisfies $\delta'(\eps)\leq C\delta(C'\eps)+C''\eps$ for some constants $C,C',C''$ depending on the presentations.  
\end{lemma}

\begin{proof}
Let $(\cM,\tau)\in \mC$ and $\phi'$ an $(\eps,\mu_{R'})$-almost homomorphism of $\langle S':R'\rangle$. Let $S=\{s_1,\ldots,s_t\}$, where $t=|S|$. For each $i\in\{1,\ldots,t\}$ there is a word $w_i\in \mathcal{F}(S')$ in the generators $S'$ and their inverses such that $s_i=w_i$ in $G$. Define $\phi(s_i)=\phi'(w_i)$ and extend $\phi$ to $\mathcal{F}(S)$ in the obvious way.

Fix an arbitrary $r\in R$. Then $r$ can be expressed as $r=s_{i_1}^{\eps_1}\cdots s_{i_\ell}^{\eps_\ell}=w_{i_1}^{\eps_1}\cdots w_{i_\ell}^{\eps_\ell}$ in $G$, where $i_1,\ldots,i_\ell\in \{1,\ldots,t\}$ and $\eps_1,\ldots,\eps_\ell\in\{\pm 1\}$. Define $w(r) = w_{i_1}^{\eps_1}\cdots w_{i_\ell}^{\eps_\ell}$ as an element of $\mathcal{F}(S')$. Since $r=1$ in $G$, also $w(r)=1$ in $G$. Moreover, we have the estimate 
\[ |w(r)| \leq |w_{i_1}|+\cdots+|w_{i_\ell}|\leq \ell \max_{i\in \{1,\ldots,t\}}|w_i|\;.\]
Let $D'=\max_{i\in \{1,\ldots,t\}}|w_i|$.

Let $\Delta'$ be the Dehn function of $\langle S':R'\rangle$. Then by definition, since $w(r)=1$ it can be written as a product 
\[ w(r)\,=\, u_1 (r'_{1})^{\eps'_1} u_1^* \cdots u_L (r'_{L})^{\eps'_L}u_L^* \]
of $L\leq \Delta'(|w(r)|) \leq \Delta'(\ell D')$ conjugates of the $r'\in R'$ and their inverses. Here, $u_i\in \mathcal{F}(S')$ and $r'_i\in R'$, for each $1\leq i \leq L$.

Using the definition of $\phi$, followed by the triangle inequality, 
\begin{align*}
\|\phi(r)-I\|_\tau &= \|\phi'(w(r))-I\|_\tau\\
&= \|\phi'( u_1 (r'_{1})^{\eps'_1}u_1^* \cdots u_L (r'_{L})^{\eps'_L} u_L^* ) - I \|_\tau\\
&\leq  \sum_{j=1}^L \|\phi'( u_j (r'_{j})^{\eps'_j}u_j^* ) - I \|_\tau\\
&=  \sum_{j=1}^L \|\phi'( r'_{j}) - I \|_\tau\\
&\leq L \max_{r'\in R'} \|\phi'(r')-1\|_\tau\;.
\end{align*}
%Let $S''=S\cup S'$ and $R''=R\cup R'$. Let $\mu_R''$ be the equal mixture of $\mu_R$ and $\mu_{R'}$.
 From the above it follows that $\phi$ is an $(\eps',\mu_R)$-almost homomorphism of $\langle S:R\rangle$ where 
\[\eps' \leq  L \eps (\min_{r'}\mu_{R'}(r'))^{-1} \;.\]
%In particular, $\phi$ is an $(\eps',\mu_R)$-almost homomorphism of $\langle S:R\rangle$. 
%Note that because $G=\langle S:R\rangle$ is stable, then $G=\langle S'':R''\rangle$ is also stable. This is
Applying the assumption, $\phi$ is $(\delta(\eps'),\mu_S)$-close to a unitary representation $g$ of $G$ on some $(\mathcal{N}, \tau^\mN)\in \mC$. 
For $s'\in S'$, write $s' = w'$ where $w'\in\mathcal{F}(S)$ is a word in the generators from $S$ and their inverses. Further replacing each generator $s_i$ from $S$ by the word $w_i$ identified at the start of the proof gives $s'=w''$ where $w''\in \mathcal{F}(S')$ is a product of the $w_i$ and their inverses.

Using the same reasoning as in the first part, it follows that 
\begin{equation}\label{eq:sg-1}
   \|\phi'(w''(s')^{-1})-I\|_\tau \leq LL' \eps (\min_{r'}\mu_{R'}(r'))^{-1}\;,
\end{equation}
where $L'=\Delta(DD')$ with $D=\max_{s'\in S'}|w'|$ and $\Delta$ the Dehn function of $\langle S:R\rangle$.
 But $\phi'(w''(s')^{-1}) = \phi'(w'')\phi'((s')^{-1})=\phi(w')\phi'(s')^{-1}$. Using Lemma~\ref{lem:g-phi} we deduce that 
\begin{align}
 \|\phi(w')-g(w')\| &\leq \sum_{i=1}^{|w'|}  \| w^* g(w'_i)w -\phi(w'_i)\|_\tau + 2\eps'\notag\\
 &\leq \delta(\eps')  (\min_{s\in S} \mu(s))^{-1} + 2\eps'\;.\label{eq:sg-2}
\end{align}
It follows from~\eqref{eq:sg-1} and~\eqref{eq:sg-2} that $\phi'$ is $\delta'(\eps)$-close to $g$, where 
\begin{align*}
\delta'(\eps) &\leq \delta(\eps')  (\min_{s\in S} \mu(s))^{-1}  + 2\eps'  +  LL' \eps (\min_{r'}\mu_R'(r'))^{-1}\\
&\leq  \delta\big(L(\min_{r'}\mu_{R'}(r'))^{-1}\eps\big)  (\min_{s\in S} \mu(s))^{-1}  + \big(2 L (\min_{r'}\mu_{R'}(r'))^{-1} +  LL'  (\min_{r'}\mu_R'(r'))^{-1}\big)\eps
\;.
%\\
%&\leq \big(LL' (\min_{s\in S} \mu(s))^{-1}(\min_{r'}\mu_R'(r'))^{-1}\big) \eps\;.
\end{align*}
\end{proof}

We now give a simple lemma that shows that homomorphisms on $\mathcal{F}(S)$ that are $(\delta,\mu_S)$ close to representations are naturally almost homomorphisms, for any presentation $\langle S:R\rangle$ and a distribution on $R$ that is naturally defined from the distribution $\mu_S$.
\tnote{This is Lemma 1 by the reviewer.}

\begin{lemma}\label{lem:l1}
  Let $G=\langle S:R\rangle$ be a finite presentation, and $\mu_R$ a distribution on $R$. For a relation $r\in R$, let $|r|$ denote its length as a word in $\mathcal{F}(S)$. Define the following measure on $S$
  \begin{equation}\label{eq:mus}
     \mu_S \,=\, \frac{1}{\sum_{r\in R} \mu_R(r)|r|^2} \sum_{r\in R} \mu_R(r) \sum_{i=1}^{|r|} \delta_{r_i}\;,
  \end{equation}
  where $\delta_{r_i}$ assigns probability $1$ to the $i$-th element of $r$, or its inverse, depending which lies in $S$. Then for any tracial von Neumann algebra $(\cM,\tau)$ and homomorphism $\phi:\mathcal{F}(S)\to \mathcal{U}(\cM)$, if $\phi$ is $(\delta,\mu_S)$ close to a representation of $G$ then $\phi$ is also an $(O(\delta),\mu_R)$-almost homomorphism. (Here the constant implicit in the $O(\cdot)$ depends linearly on $\Es{r\sim \mu_R}|r|^2$.)
  \end{lemma}

  \begin{proof}
    Let $\phi:\mathcal{F}(S)\to \mathcal{U}(\cM)$ be as in the lemma, and let $g$ be a $(\delta,\mu_S)$ close representation of $G$.
   Using the triangle inequality and $g(r)=1_\mN$ and then Lemma~\ref{lem:g-phi}, 
  \begin{align*}
  \Es{r\sim \mu_R}\|\phi(r)-\Id\|_\tau &\leq \Es{r\sim \mu_R}\|\phi(r)-w^*g(r)w\|_\tau + \|w^*w-I_\mM\|_\tau\\
  &\leq \Es{r\sim \mu_R}\sum_{i=1}^{|r|}\| w^* g(r_i)w -\phi(r_i)\|_\tau + \|w^*w-I_\mM\|_\tau +2\delta \\
  &=  \sum_{s\in S} \sum_{r\in R} \mu(R) \sum_{i=1}^{|r|} 1_{s\in \{r_i,r_i^{-1}\}} \| w^* g(r_i)w -\phi(r_i)\|_\tau +   \|w^*w-I_\mM\|_\tau +2\delta\\
  &\leq \Big(\sum_{r\in R} \mu_R(r)|r|^2\Big)\Es{s\sim \mu_S }5 \| w^* g(s)w -\phi(s)\|_\tau +    \|w^*w-I_\mM\|_\tau +2\delta\\
  &\leq 8\delta \Big(\sum_{r\in R} \mu_R(r)|r|^2\Big)\;,
  \end{align*}
  where the factor of $5$ that appeared on the third line accounts for the application of~\eqref{eq:switch-inv} whenever necessary, and for the last line we used the fact that  $\Es{r\sim \mu_R}|r|^2\geq 1$. 
  \end{proof}

  \tnote{changed remark below based on referee comment}
\begin{remark}\label{rk-mus}
  Throughout this text, whenever $\mu_R$ is specified by not $\mu_S$ we mean that $\mu_S$ is chosen as the induced distribution defined in~\eqref{eq:mus}.
\end{remark}

Definition~\ref{def:eff-stab} specifies what it means for a presentation to be stable, but not when the presentation is \emph{efficiently} stable. 
As mentioned in the introduction, the goal is to optimise the tradeoff between the encoding length of the presentation and the resulting modulus of stability. We now choose a complexity measurement for presentations:
\begin{definition}[Length of a presentation]\label{def:length_of_pres}
    Let $G=\langle S\colon R\rangle$ be a finite presentation. The \emph{length} of the presentation  will be  $$\ell(G)=|S|+ \sum_{r\in R}|r|\;.$$ 
\end{definition}

@@

As was shortly reviewed in the introduction, and specifically in the related works subsection, many results about various notions of stability are known.  Here, we give two results that are most relevant to our work. First of all, 
for a finite group $G$ we can always write $G=\langle S:R\rangle$ where $S = G$ and $R=\{ g\cdot h \cdot (gh)^{-1} =e \}$. We refer to this presentation as the \emph{multiplication table presentation}. If we let $\mu_S$ and $\mu_R$ be the uniform distribution on $S$ and on $R$ respectively then Definition~\ref{def:eff-stab} reduces to a widely used notion of \emph{flexible (Hilbert-Schmidt) stability}. In particular, for finite groups the following result is known~\cite{gowers2017inverse,de2019operator}. We adopt the formulation from~\cite[Theorem 1.4]{de2022spectral}.

@@

\begin{remark}\label{rk:linear-modulus}\tnote{adding:}
  We remark that, using Lemma~\ref{lem:stab-group}, Theorem~\ref{thm:gh} implies that any finite presentation $G=\langle S:R\rangle$ of a finite group $G$ is $(O(\eps),\mu_S,\mu_R,\mathcal{C})$-stable, where $\mu_R$ is any distribution with full support and $\mu_S$ is e.g.\ the distribution induced from $\mu_R$ as in Lemma~\ref{lem:l1}. However, the constant implicit in the $O(\eps)$ modulus of stability may of course depend on the presentation, and this will be our focus later on. 
\end{remark}

@@

\begin{equation}\label{eq:z2-efficient}
 \Z_2^k = \langle x_1,\ldots,x_k : [x_i,x_j]=e, x_i^2=e \; \forall i\neq j \rangle\;,
\end{equation}
where $[x_i,x_j]=x_ix_jx_i^{-1}x_j^{-1}$ is the group commutator. 
Its length is $O(k^2)$, which is polylogarithmic in the group size instead compared to the polynomial length of the multiplication table presentation. By \cite{hadwin2018stability}, every presentation of a finite group has \textbf{some} modulus of stability.
The following two lemmas show that, though the presentation \eqref{eq:z2-efficient} has a linear modulus of stability (as expected, see Remark~\ref{rk:linear-modulus}), it deteriorates by a constant factor $k$. In some sense, these two examples are on opposite sides of the efficient stability tradeoof: \eqref{eq:z2-efficient} is short but has a bad modulus of stability, while the multiplication table presentation is very long, while having an essentially optimal modulus of stability. As we keep recalling, the goal of this paper is to provide an example of a somewhat short presentation of $\Z_2^k$, which has a good enough modulus of stability to deduce the main technical results needed for $\MIP^*=\RE$~\cite{ji2020mip}.

@@

As one would expect from Lemma~\ref{lem:stab-group}, the dependence of $\delta_k$ on $k$ depends on both the diameter of $\Z_2^k$ for the presentation~\eqref{eq:z2-efficient}, and certain values of its \emph{Dehn function} --- namely, the minimal volume of a Van Kampen diagram with perimeters of length $3$. By expressing each element of $\Z_2^k$ as a product of generators in the natural way, and by applying Corollary~\ref{cor:lin-test}, it is possible to show that $\delta_k=O(k^2\eps)$ in Lemma~\ref{lem:eff-z2}. With more work one can get $\delta_k=O(k\eps)$, see~\cite[Theorem 3.2]{chao2017overlapping}. This turns out to be  tight, as the next lemma shows.

@@

Lemma~\ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k} implies that the modulus of stability of \eqref{eq:z2-efficient} is $\Omega(k\eps)$ whenever $\eps\leq\frac{1}{k-1}$. This is because, for any $0<\eps<\nicefrac{1}{k-1}$, one can add trivial representations of $\Z_2^k$ to copies of the approximate representation described in Lemma~\ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k} (choosing $c=\nicefrac{k}{2}$), such that the resulting map is an $(\eps,\mu_R)$-approximate representation which is at least $(\nicefrac{(k-1)\eps}{32},\mu_S)$-away from any actual representation of $\Z_2^k$. We record this observation as the following corollary.

\begin{corollary}\label{cor:delta-z2k}
The presentation of $\Z_2^k$ given in~\eqref{eq:z2-efficient}, together with the distributions $\mu_R$ and $\mu_S$ specified in Lemma~\ref{lem:eff-z2}, has a modulus of stability  $\delta_k(\eps)$ with respect to the class $\mathcal{C}$ of tracial von Neumann algebras that satisfies
\[ \frac{1}{C} \min(k\eps,1)\,\leq\,\delta_k(\eps)\,\leq\, C\min(k\eps,1)\;,\]
where $C>0$ is some constant.
\end{corollary}

The proof of Lemma~\ref{lem:lower_bound_on_stability_rate_standard_presentation_Z_2^k} is given in Appendix \ref{appendix:lower_bounds}, which also includes some discussion on the $L^\infty$ variant of it.

@@

\begin{proof}
This follows from~\cite{chao2017overlapping} and Theorem~\ref{thm:dls-gap}, in a similar manner to our proof sketch of Corollary \ref{cor:Pauli-brading_is_stable}.
\end{proof}

@@

\begin{proof}
The group defined by the right-hand side is obviously abelian and a $2$-group, so it is of the form $\Z_2^{k'}$ for some $k'$. In fact, it is isomorphic to the quotient of $\Z_2^n$ by the subgroup generated by the $\prod_{1\leq j \leq n} x_j^{h_{ij}}$. So it is $\Z_2^k$ where $k=n-\dim\textrm{im}\ h = \dim\ker h=\dim \code$.  
\end{proof}

@@

\begin{definition}[$r$-local linear tester]\label{def:code-test}
Let $\code$ be an $[n,k,d]_2$ linear code and $r\in \N$.
An \emph{$r$-local linear tester for $\code$} is a pair $M = (h,\nu)$ where $h \in \F_2^{m \times n}$ is a parity check matrix for $\code$, whose every row has Hamming weight at most $r$, and $\nu$ is a distribution over $\{1,\ldots,m\}$. 
\end{definition}

@@

A parity check matrix $h_{\RM} \in \F_q^{S \times q^m}$ for $\code_{\RM}$, where $S = q^m \times m \times q$, is as follows. Identify the rows with triples $(u,j,t) \in \F_q^m \times \{1,\ldots,m\} \times \F_q$. The $(u,j,t)$-th row of $h_{\RM}$ is the vector in $\F_q^m$ that for $i \in \{0,\ldots,d\}$ has the value $\alpha_{u,v,i}$ for $v = u + t e_j$ in the coordinate indexed by $u + t_i e_j$, the value $-1$ in the coordinate indexed by $u + te_j$, and the value $0$ everywhere else.

@@

 The Reed-Muller code from the previous section is defined over $\F_q$, for $q=2^t$ a power of $2$. We can transform any $q$-ary code, for $q=2^t$, into a binary code using the idea of \emph{code composition} which we now describe.

@@

%Let $\tr(\cdot):\F_q\to\F_2$ denote the trace over $\F_2$, wich is defined by $\tr(x) = \sum_{j=0}^{t-1} x^{2^i}$ for $x\in \F_q$. Fix a  self-dual basis $\{e_1,\ldots,e_t\}$  of $\F_q$ over $\F_2$. (A self-dual basis $\{e_1,\ldots,e_t\}$ is one which satisfies $\tr(e_ie_j)=\delta_{ij}$ for all $1\leq i,j\leq t$.)
%Let $\kappa: \F_q \to \F_2^t$ denote the invertible linear map such that $\kappa(a)$ is the vector of coefficients of $a\in \F_q$ in the basis $\{e_1,\ldots,e_t\}$, i.e.\ $\kappa(a)_i=\tr(ae_i)$ for $i\in\{1,\ldots,t\}$.

%We extend $\kappa$ and its inverse $\kappa^{-1}$ to vectors over $\F_q$ coordinate-wise. Because we chose a self-dual basis for the binary representation, the trace satisfies $\tr(ab)=\kappa(a)\cdot\kappa(b)$  where the right-hand side means the $\F_2$-inner product of the vectors $\kappa(a), \kappa(b)$.

\tnote{edited next two paragraph following referee comments:}
Let $q=2^t$ and $\code$ be an $[n,k,d]_q$ linear code. We define an $[qn,tk,d']_2$ linear code as follows. First introduce the $\F_2$-linear map $\varphi:\F_q\to\F_2^q$ defined by $\varphi(x)=(\tr(xa))_{a\in \F_q}$. Here  $\tr(\cdot):\F_q\to\F_2$ denotes the trace over $\F_2$, wich is defined by $\tr(x) = \sum_{j=0}^{t-1} x^{2^i}$ for $x\in \F_q$. The code $\code'$ is defined as the linear span, over $\F_2$, of all $(\varphi(x_1),\ldots,\varphi(x_n))$ for $(x_1,\ldots,x_n)\in \code$. It is apparent from the definition that this code has dimension $tk$, the dimension of $\code$ over $\F_2$, and length $qn$. Furthermore, it is not hard to verify that it has distance $d'\geq dq/2$.

The reason that this construction is referred to as ``composition'' is the following. Observe that the linear span, over $\F_2$, of all $\varphi(x)$, $x\in \F_q$, is (up to reordering of the indices) identical to the Hadamard code $\code_\Had$ over $\F_2^t$ introduced at the end of Section~\ref{sec:pres-code}. Thus each $\varphi(x)$ can be interpreted as an ``encoding'' of the element $x\in \F_q$ (when seen as an element of $\F_2^t$ by fixing a basis); and  $(\varphi(x_1),\ldots,\varphi(x_n))$ is the successive encoding of an element $m\in\F_2^k$ through $\code$ (to obtain $(x_1,\ldots,x_n))$ and then $\code_\Had$.

%
%Let $\code_{\Had}$ be the Hadamard code . 
%Let $\code'$ be the $[qn,tk,d']$ linear code over $\F_2$ defined as follows. First, define a linear map $Enc: \F_2^{tk}\to \F_2^{qn}$ as follows. For $a\in \F_2^{tk}$, interpret $a$ as an element of $(\F_2^t)^k$ in some arbitrary way. Map $a\mapsto a'=\kappa^{-1}(a) \in \F_q^{k}$. Then encode $a'$ to $b'=\code(a')\in \F_q^n$. Finally, return $b=\code_\Had(\kappa(b'))\in(\F_2^q)^n$, where $\code_\Had$ is applied component-wise. Using that $\code_\Had$ has distance $q/2$, it is easy to verify that this code has distance $d'\geq dq/2$.

Given an $r$-local tester $M=(h,\nu)$ for $\code$, there is a natural $\max(r,3)$-local tester $M'$ for $\code'$ which can be described as follows. Index coordinates of $\code'$ by pairs $(i,\alpha)\in [n]\times\F_2^t$, fixing a bijection between $[qn]$ and $[n]\times \F_2^t$.  We describe an $\max(r,3)$-local tester $M' = (h',\nu')$ for $\code'$. Informally, $h'$ contains two type of checks. First, the $A$-checks consist of the repetition of $n$ copies of the checks for the Hadamard code, one for each Hadamard-code encoding $\varphi(x)$ of an $\F_q$-symbol $x$ from $\code$. Second, the $B$-checks implement the checks of $\code$ specified by $M$, directly on the Hadamard encoding.
